Using Monte Carlo to estimate the effects of NHL rule changes

### 2020-12-10 AltSeasonSimulation.Rmd
Add files via upload
Simulation for 2017-2018 season, copy of simulation.rmd with some minor modifications

---
title: "Simulations"
output: pdf_document
---

@@

# Read and filter the data
```{r}
data <- read_csv("Data/game.csv")
data <- data %>% filter(data$season == 20132014, data$home_team_id %in% c(seq(1,30), 52), data$away_team_id %in% c(seq(1,30), 52), data$type == 'R')
data$home_team_id[data$home_team_id == 52] <- 11
data$away_team_id[data$away_team_id == 52] <- 11
```

# Simulating a season
```{r}
simulateSeason <- function() {
  # Draw a "team quality" value for all teams.
  teamQuality <- rnorm(30, 0, sigma)

@@

```{r}
# Get the sum of the points for each team. The return vector is sorted by team number value.
calcTotalPoints <- function(season){
  totalPoints <- rep(0, 30)
  for (game in 1:length(season[[1]])){
    totalPoints[season[[2]][game]] = totalPoints[season[[2]][game]] + season[[7]][game]
    totalPoints[season[[3]][game]] = totalPoints[season[[3]][game]] + season[[8]][game]
  }
  returnVec = data.frame("id" = seq(1,30), totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))

@@

returnNewWestern <- function(points) {
  central = points[points$id %in% c(11, 16, 18, 19, 21, 25, 30),]
  pacific = points[points$id %in% c(20, 22, 23, 24, 26:28),]

@@

  wildcards = rbind(central[4:7,],pacific[4:7,])
  wildcards = wildcards[order(-wildcards$totalPoints),]

@@

  for (i in c(11, 16, 18:28, 30)){
    western[indexW, 1] = points[i, 1]
    western[indexW, 2] = points[i, 2]
    indexW = indexW + 1
  }
  return(c(eastern, western))
}
```

@@

for (i in 1:nloop) {
  season <- simulateSeason()
  points <- calcTotalPoints(season)
  oldEast <- returnOldEastern(points)
  oldWest <- returnOldWestern(points)
  newEast <- returnNewEastern(points)
  newWest <- returnNewWestern(points)
  oldGap[i] <- oldEast[8,2] - oldWest[8,2]
  newGap[i] <- newEast[8,2] - newWest[8,2]
  allTeams <- rbind(newEast, newWest)
  allTeams <- allTeams[order(-allTeams$totalPoints),]
  seeds <- data.frame(seedPlace = seq(1,30), points = allTeams$totalPoints)
  seedPoints <- rbind(seedPoints, seeds)
}

>>>>

title: "Simulations"
output: pdf_document
---

@@

# Read and filter the data
```{r}
data <- read_csv("Data/game.csv")

data <- data %>% filter(data$season == 20172018, data$home_team_id %in% c(seq(1,30), 52:54), data$away_team_id %in% c(seq(1,30), 52:54), data$type == 'R')

data$home_team_id[data$home_team_id == 52] <- 11
data$away_team_id[data$away_team_id == 52] <- 11
data$home_team_id[data$home_team_id == 53] <- 27
data$away_team_id[data$away_team_id == 53] <- 27
data$home_team_id[data$home_team_id == 54] <- 31
data$away_team_id[data$away_team_id == 54] <- 31
```

# Simulating a season
```{r}
simulateSeason <- function() {
  # Draw a "team quality" value for all teams.
  teamQuality <- rnorm(31, 0, sigma)

@@

```{r}
# Get the sum of the points for each team. The return vector is sorted by team number value.
calcTotalPoints <- function(season){
  totalPoints <- rep(0, 31)
  for (game in 1:length(season[[1]])){
    totalPoints[season[[2]][game]] = totalPoints[season[[2]][game]] + season[[7]][game]
    totalPoints[season[[3]][game]] = totalPoints[season[[3]][game]] + season[[8]][game]
  }
  returnVec = data.frame("id" = seq(1,31), totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild","Vegas Golden Knights"))

@@

returnNewWestern <- function(points) {
  central = points[points$id %in% c(11, 16, 18, 19, 21, 25, 30),]
  pacific = points[points$id %in% c(20, 22, 23, 24, 26:28,31),]

@@

  wildcards = rbind(central[4:7,],pacific[4:8,])
  wildcards = wildcards[order(-wildcards$totalPoints),]

@@

  for (i in c(11, 16, 18:28, 30:31)){
    western[indexW, 1] = points[i, 1]
    western[indexW, 2] = points[i, 2]
    indexW = indexW + 1
  }
  return(c(eastern, western))
}
```

@@

for (i in 1:nloop) {
  season <- simulateSeason()
  points <- calcTotalPoints(season)
  oldEast <- returnOldEastern(points)
  oldWest <- returnOldWestern(points)
  newEast <- returnNewEastern(points)
  newWest <- returnNewWestern(points)
  oldGap[i] <- oldEast[8,2] - oldWest[8,2]
  newGap[i] <- newEast[8,2] - newWest[8,2]
  allTeams <- rbind(newEast, newWest)
  allTeams <- allTeams[order(-allTeams$totalPoints),]
  seeds <- data.frame(seedPlace = seq(1,31), points = allTeams$totalPoints)
  seedPoints <- rbind(seedPoints, seeds)
}
